MATH/COMP 562 ASSIGNMENT 3

### 2023-06-24 assignment-3/article.tex
Solve random variable questions
Write up algorithmic stability

@@

        \item Conclude that the Ridge Regression problem is replace one stable (ROS) and give the rate as a function of $\lambda$, $m$.

        \[\] % Space to write solution
    \end{enumerate}

    \[\] % Space to write solution
\end{enumerate}

>>>>

\begin{solution}
  \begin{enumerate}
    \item
      \begin{enumerate}
        \item $E[X] = \alpha w + (1-\alpha)u$ therefore $E[X]^2 = ||\alpha w + (1-\alpha)u||^2$
        \item $E[X^2] = \alpha||w||^2 + (1-\alpha) ||u||^2$
        \item $\begin{aligned}[t]
          Var[X] &= E[X^2] - E[X]^2 \\
          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - ||\alpha w + (1-\alpha)u||^2 \\
          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - (\alpha w + (1-\alpha)u)^T(\alpha w + (1-\alpha)u) \\
          &= \alpha||w||^2 + (1-\alpha) ||u||^2 - \left(\alpha^2||w||^2 -2\alpha(1-\alpha)(u\cdot w) + (1-\alpha)^2||u||^2\right) \\
          &= ||w||^2(\alpha -\alpha^2)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2(1-\alpha-(1-\alpha)^2) \\
          &= ||w||^2\alpha(1 -\alpha)+2\alpha(1-\alpha)(u\cdot w) + ||u||^2\alpha(1 - \alpha) \\
          &= \alpha(1-\alpha)\left(||w||^2 - 2(u\cdot w) + ||u||^2\right) \\
          &= \alpha(1-\alpha)||w-u||^2
          \end{aligned}$
      \end{enumerate}
    \item $\begin{aligned}[t]
      E[f(X)] - f(E[X]) &= \alpha f(w) + (1-\alpha)f(u) - f(\alpha w + (1-\alpha)u) \\
      &\geq \alpha f(w) + (1-\alpha)f(u) - \alpha f(w) -(1-\alpha)f(u) + \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 \\
      &\geq \frac{\lambda}{2}\alpha(1-\alpha)||w-u||^2 = \frac{\lambda}{2}Var[x]
      \end{aligned}$
  \end{enumerate}
\end{solution}

@@

        \item Conclude that the Ridge Regression problem is replace one stable (ROS) and give the rate as a function of $\lambda$, $m$.
    \end{enumerate}
\end{problem}

\begin{solution}
  \begin{enumerate}
    \item $g$ is $\lambda$-strongly convex by the proof in exercise 3.5(c)
    \item Since $g$ is $\lambda$-strongly convex, we apply the inequality setting $\alpha=0.5$): $$g(\frac{1}{2} w_1 + \frac{1}{2}w_2, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2} g(w_1, S_1) + \frac{1}{2}g(w_2, S_1)$$

      Because $w_1$ minimises $g$ at $S_1$, we have that $g(w_1, S_1) \leq g(\frac{1}{2} w_1 + \frac{1}{2} w_2, S_1)$. So we can substitute $g(w_1, S_1)$ in the left hand side in the following way: $$g(w_1, S_1) + \frac{\lambda}{4}||w_1-w_2||^2 \leq \frac{1}{2}g(w_1, s_1) + \frac{1}{2}g(w_2, S_1)$$

      Subtracting $g(w_1, S_1)$ from both sides of the inequality, and then multiplying both sides by 2 gives the first desired inequality:
      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2,S_1)-g(w_1, S_1)$$
      We can obtain the other inequality by the same argument, or via symmetry.
      $$\frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1,S_2)-g(w_2, S_2)$$
    \item We start by rewriting the inequalities from the preceding exercise.
      \begin{align*}
        \frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_2, z_1)-g(w_1, z_1) \\
        \frac{\lambda}{2}||w_1-w_2||^2 \leq g(w_1, z_2)-g(w_2, z_2)
      \end{align*}
      Then we add both inequalities together to obtain:
      \begin{align*}
        \lambda ||w_1-w_2||^2 &\leq (g(w_2,z_1)-g(w_1,z_1)+g(w_1,z_2)-g(w_2,z_2)) \\
        \hphantom{\lambda ||w_1-w_2||^2} &\leq (f(w_2,z_1) + ||w_2||^2 -f(w_1,z_1) - ||w_2||^2 \\
        \hphantom{\lambda ||w_1-w_2||^2}&+f(w_1,z_2) + ||w_1||^2-f(w_2,z_2) - ||w_1||^2 \\
        &\leq f(w_2,z_1)-f(w_1,z_1)+f(w_1,z_2)-f(w_2,z_2) \\
        &\leq \frac{1}{m} (\ell(w_2,z_1)-\ell(w_1,z_1)+\ell(w_1,z_2)-\ell(w_2,z_2))
      \end{align*}
      where the last inequality comes from the fact that $f(w_i,S_i) = \frac{1}{m}\ell(w_i, z_i)$.
      \item We begin by calculating an upper bound for $\ell(w,z)$: \begin{align*}
        \ell(w, z) &= (w \cdot x - y)^2 \\
        &= ||w\cdot x||^2 - 2 (w \cdot x) \cdot y + ||y||^2 \\
        &\leq ||w||^2 ||x||^2 - 2||w|| ||x|| ||y|| + ||y||^2 \\
        &\leq ||w||^2 ||x||^2 + ||y||^2 \\
        &\leq C_w^2C_x^2 + C_y^2
        \end{align*}

        Now we use the hint:
        \begin{align*}
          \ell(w_2,z_1) - \ell(w_1,z_1) &= (w_2\cdot x - y)^2 - (w_1\cdot x - y)^2 \\
          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= (w_2\cdot x - y - w_1\cdot x + y)(w_2\cdot x - y + w_1\cdot x - y) \\
          \hphantom{\ell(w_2,z_1) - \ell(w_1,z_1)} &= ((w_2 - w_1)\cdot x)((w_1 + w_2)\cdot x - 2y)
        \end{align*}
  \end{enumerate}
\end{solution}
